---
title: "STEM students perception of bioinformatics"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

```{r}
library(RColorBrewer) 
library(wordcloud2)
library(ggplot2)
```
 
# Import data 
Read in data and split dataframe between STEM and Bioinformatics students
```{r}
survey<-read.csv2("../../data/BioinformaticsSurvey2022.csv",sep=',') 
survey_bioinf<-survey[survey$Is.bioinformatics.the.main.focus.of.your.studies.=='Yes',c(1:7,26:77)]
survey_STEM<-survey[survey$Is.bioinformatics.the.main.focus.of.your.studies.=='No',c(1:25,74:77)]
survey_STEM
```

# Word cloud
Create a word-cloud with the terms used by STEM students to describe bioinformatics
```{r}
# extract words 
words<-unlist(strsplit(survey$How.would.you.describe.bioinformatics.in.3.words..Please..write.3.words.separated.by.a.comma.in.small.letters..without.using.any.special.characters.,',',fixed=T))

# preprocess
words<-trimws(words)
words<-tolower(words)
#table(words) 

# correct words
tmp<-words[28]
words<-words[-28]
words<-c(words,trimws(unlist(strsplit(tmp,'/'))))
words<-words[words!="the use of a informatica tool to understand anything bio related"] 
words[grep('machine',words)]<-'machine-learning'
words[grep('single',words)]<-'single-cell'
words[grep('big',words)]<-'big-data'
words[grep('rapid',words)]<-'rapid'

words<-unlist(strsplit(words,' '))
words[grep('cha',words)]<-'challenging'
words[grep('stat',words)]<-'statistics'
words[grep('useful',words)]<-'useful'
words[grep('utile',words)]<-'useful'
words[grep('model',words)]<-'modeling'
words[grep('analisis',words)]<-'analysis'
words[grep('anal',words)]<-'analysis'
words[grep('tool',words)]<-'tools'
words[grep('programs',words)]<-'tools' 
words[grep('seq',words)]<-'sequencing'
words[grep('algorithm',words)]<-'algorithm'
words[grep('science',words)]<-'science'
words[grep('powerfull',words)]<-'powerful'
words[grep('confusa',words)]<-'confusing'
words[grep('programmazione',words)]<-'programming'
words[grep('moderna',words)]<-'modern'
words[grep('pred',words)]<-'prediction'
words[grep('informatic',words)]<-'informatics'
words[grep('precis',words)]<-'precise' 
words[grep('sistemi',words)]<-'systems' 
words[grep('pratic',words)]<-'practical' 
words[grep('found',words)]<-'fundamental' 
words[grep('intere',words)]<-'interesting'
words[grep('intesting',words)]<-'interesting'
words[grep('biol',words)]<-'biology'
words[grep('visulization',words)]<-'visualization'
words[grep('helpful',words)]<-'helpful'
words[grep('innov',words)]<-'innovative'
words[grep('complex',words)]<-'complexity'
words[grep('fondamental',words)]<-'fundamental'
words[grep('computation',words)]<-'computation'
words[grep('database',words)]<-'databases'
words[grep('comunication',words)]<-'communication'
words[grep('genetic',words)]<-'genetics'
words[grep('computing',words)]<-'computation'
words[grep('geomes',words)]<-'genomes'
words[grep('indispensabile',words)]<-'indispensable'
words[words%in%'omic']<-'omics'
words<-words[!words%in%c('a','in','very','the','i','and','of','to','for','know','dont',"don’t")]
```

```{r}
#table(words) 
wordcloud2(data.frame(table(words)),shape='circle',color='random-dark',shuffle=F,size=1.2) 
```
 
# STEM student's knowledge on bioinformatics (stratified by degree ares)
We plot for each degree area whether the students heard about bioinformatics
```{r}  
df<-survey_STEM[,c("What.is.your.current.degree.area..Please.select.the.closest.answer.that.applies..",
                         "Have.you.ever.heard.about.bioinformatics.")]
colnames(df)<-c('Degree','Ever heard about bioinformatics')
df<-as.data.frame(table(df))
ggplot(df,aes(fill=Ever.heard.about.bioinformatics,
              y=Freq,
              x=Degree)) +geom_bar(position='dodge',stat='identity')   +theme(axis.text.x = element_text(angle = 90))
```

# STEM student's bioinformatics courses (stratified by degree area)
We plot for each degree area whether the students ever had a bioinformatics course
```{r}
df<-survey_STEM[,c("What.is.your.current.degree.area..Please.select.the.closest.answer.that.applies..",
                         "Have.you.ever.taken.a.course.in.computational.biology.bioinformatics.")]
colnames(df)<-c('Degree','Ever had a bioinformatics course')
df<-as.data.frame(table(df))
ggplot(df,aes(fill=Ever.had.a.bioinformatics.course,
              y=Freq,
              x=Degree)) +geom_bar(position='dodge',stat='identity')   +theme(axis.text.x = element_text(angle = 90))

```

# Perception of workplace between STEM and bioinformaticians
We investigate the relationship between the perception of STEM students of where a bioinformatician works and where bioinformatics students see themwelves working in the future
```{r}
workplace<-c('Academia','Industry (R&D)', 'Data Analyst','IRCSS','Communication','Teaching','Entepreneurship','Editorial')
STEM_indexes<-13:20
BIOINFO_indexes<-STEM_indexes+33

STEM_workplace<-survey_STEM[,STEM_indexes]
colnames(STEM_workplace)<-workplace

BIOINFO_workplace<-survey_bioinf[,BIOINFO_indexes]
colnames(BIOINFO_workplace)<-workplace
```  

Plot the average score the two categories gave to different workplaces
```{r}
avg_stem_workplace<-colMeans(STEM_workplace) 
avg_bioinfo_workplace<-colMeans(BIOINFO_workplace) 
sprintf('Correlation of average score: %f', cor(avg_stem_workplace,avg_bioinfo_workplace))

plot(x=rep(c(1:8),2),y=c(avg_stem_workplace,avg_bioinfo_workplace), xlab = '', ylab='Score', ylim=c(1,5),axes = FALSE, col=c(rep('red',8),rep('blue',8)),pch=20)#+plot(avg_bioinfo_workplace,  col='blue', pch=20)
axis(2) 
box()
axis(1, at=1:8, labels = workplace, las=2, cex.axis=0.6)  
legend('topright',legend=c('stem','bioinformatics'),col=c('red','blue'),pch=20)
```

For each workplace, we compute how many people gave scores from 1-5 in the two categories (stem/bioinformatics) and compute the correlation between STEM and bioinformatics students in each workplace (NOT SURE THIS MAKES SENSE. Should it be normalized? Does it make sense to compute correlation between 5 data points?)
```{r}
correlation<-c()
heatmap_stem<-c()
heatmap_bioinfo<-c()
for (i in 1:length(workplace)){ 
  s<-c()
  b<-c()
  for (n in 1:5){
  s[[n]]<-length(which(STEM_workplace[,i]==n))
  b[[n]]<-length(which(BIOINFO_workplace[,i]==n))
  }
  correlation[[workplace[i]]]<-cor(unlist(s),unlist(b))
  heatmap_stem<-rbind(heatmap_stem,unlist(s))
  heatmap_bioinfo<-rbind(heatmap_bioinfo,unlist(b))
}

rownames(heatmap_bioinfo)<-workplace
rownames(heatmap_stem)<-workplace

colnames(heatmap_bioinfo)<-c(1:5)
colnames(heatmap_stem)<-c(1:5)

correlation
```
Plot how many people gave scores from 1-5 in the two categories
```{r} 
heatmap(heatmap_bioinfo, Rowv = NA, Colv = NA, main = 'Bioinformatics students')
heatmap(heatmap_stem, Rowv = NA, Colv = NA, main = 'STEM students')
 
```

# Perception of bioinformatician activities
We visualize the average score given by STEM students to different activities
```{r}
activity<-c('Drug-development','Diagnostics', 'Disease.characterization','Software Development','Data Analysis')
activity_indexes<-21:25
STEM_activity<-survey_STEM[,activity_indexes]
colnames(STEM_activity)<-activity
```

```{r}
# average
avg_stem_activity<-colMeans(STEM_activity) 
plot(avg_stem_activity, xlab = '', ylab='Score', ylim=c(1,5),axes = FALSE, col='red',pch=20)
axis(2) 
box()
axis(1, at=1:5, labels = activity, las=2, cex.axis=0.4)   
```

We visualize the number of votes given to each score and activity
```{r} 
heatmap_activity<-c() 
for (i in 1:length(activity)){ 
  s<-c() 
  for (n in 1:5){
    s[[n]]<-length(which(STEM_activity[,i]==n)) 
  } 
  heatmap_activity<-rbind(heatmap_activity,unlist(s)) 
}

rownames(heatmap_activity)<-activity 
colnames(heatmap_activity)<-c(1:5) 

heatmap(heatmap_activity, Rowv = NA, Colv = NA, main = 'STEM students perception of bioinformatics activities', cexRow = 0.8)
```
