---
title: "ISCB-RSG Italy: Bioinformatic perception survey, 'Who are you?'section"
author: "ISCB RSG Italy Committee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    code_folding: "hide"
    toc: true
    toc_float: true
    theme: united
---


```{r, setup, hide = TRUE, echo=FALSE, results = FALSE, warning=FALSE}
library(plyr)
library(reticulate)
library(tidyverse)
library(rcartocolor)
# To run this .rmd file:
#export PATH=${PATH}:/cm/shared/apps/R/deps/rstudio/bin/pandoc
#file="./Who_are_you_section_analysis.Rmd"
#module load R
#Rscript -e 'rmarkdown::render("'$file'")'
```
# Take the columns for 'Who are you ?' section

```{r,hide = TRUE}
survey <- read.table("../../data/BioinformaticsSurvey2022.csv",sep=",",header=TRUE)
#take only the questions in the 'Who are you ?' section:
way <- survey %>% select("How.old.are.you.","What.are.your.pronouns.","Where.are.you.from.",
                         "Where.in.Italy.are.you.studying.did.you.study.",          "Are.you.an.off.site.student.","What.is.your.current.position.","Is.bioinformatics.the.main.focus.of.your.studies.",
"What.is.your.current.degree.area..Please.select.the.closest.answer.that.applies..")
#columns as factor:
way <- way %>% mutate_if(is.character,as.factor)
#rename to shorten levels with the Italian areas:
way$Where.are.you.from. <- revalue(way$Where.are.you.from., c("Italian Islands (Sicilia, Sardegna)"="Italian Islands",
"Central Italy (Toscana, Umbria, Lazio, Marche)"="Central Italy",
"Northern-East Italy (Trentino-Alto Adige, Veneto, Friuli-Venezia Giulia, Emilia-Romagna)"="Northern-East Italy",
"Northern-West Italy (Valle d'Aosta, Liguria, Lombardia, Piemonte)"="Northern-West Italy",
"Southern Italy (Abruzzo, Molise, Campania, Puglia, Basilicata, Calabria)"="Southern Italy"))

way$Where.in.Italy.are.you.studying.did.you.study. <- revalue(way$Where.in.Italy.are.you.studying.did.you.study., c("Italian Islands (Sicilia, Sardegna)"="Italian Islands",
"Central Italy (Toscana, Umbria, Lazio, Marche)"="Central Italy",
"Northern-East Italy (Trentino-Alto Adige, Veneto, Friuli-Venezia Giulia, Emilia-Romagna)"="Northern-East Italy",
"Northern-West Italy (Valle d'Aosta, Liguria, Lombardia, Piemonte)"="Northern-West Italy",
"Southern Italy (Abruzzo, Molise, Campania, Puglia, Basilicata, Calabria)"="Southern Italy"))

way$How.old.are.you. <- revalue(way$How.old.are.you., c("Under 18"=">18",
                                                        "Over 35"="35>"))

#way$What.is.your.current.position. <- revalue(way$What.is.your.current.position., c(
#  "Bachelor’s student"="Bachelor student",
#  "Employed in other academic positions (research assistant, facility technician, …)"="Employed in other academic positions",
#  "Employed outside academia (company, start-up, spin-off, …)"="Employed outside academia",
#  "PhD student (either in academia or industry)"="PhD student"))
```

# Age and gender

The 56.70% of the participants are in the 23-26 age group, followed by a 24.52% of participants belonged to 19-22 age group.
<br>
More than half of the participants declare to self-identify with the pronouns 'she/her' (57.09%); 39.85% self-identify with the pronouns 'he/him'.
```{r,hide = TRUE}
prop.table(table(way$How.old.are.you.))*100
prop.table(table(way$What.are.your.pronouns.))*100
```

# Where are you from?

```{r,hide = TRUE}
ggplot(data = way,
       aes(x = Where.are.you.from.,
           fill = Where.are.you.from.)) + 
  geom_bar(aes(y = (..count..)/sum(..count..)), show.legend=FALSE) +
  scale_y_continuous(labels=scales::percent) +
  labs(x = "Continent or Italian areas", 
       y = "Relative frequency", 
       title  = "Where are you from ?") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  scale_fill_manual(values=c("#798234","#a3ad62","#d0d3a2","#fdfbe4","#f0c6c3","#df91a3","#d46780","#d1eeea"))
ggsave("Where_are_you_from.png")

ggplot(data = way,
       aes(x = Where.are.you.from., 
           fill = Are.you.an.off.site.student.)) + 
  geom_bar(position = "dodge") +
  labs(x = "Italian areas",
       title  = "'Where are you from?' vs 'Are you an off-site student?'",
       fill = "Off-site student status") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
   scale_fill_manual(values=c("#66C5CC","#DCB0F2"))
ggsave("Where_are_you_from_by_offsite_student.png")
```

# 'Where in Italy are you studying/did you study ?' with 'Where are you from?'

```{r,hide = TRUE}
ggplot(data = way,
       aes(x = Where.in.Italy.are.you.studying.did.you.study., 
           fill = Where.are.you.from.)) + 
  geom_bar(position = "dodge") +
  labs(x = "Italian areas",
       title  = "'Where in Italy are you studying/did you study?' vs 'Where are you from?'",
       fill = "Continent or Italian areas of origin") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
   scale_fill_manual(values=c("#798234","#a3ad62","#d0d3a2","#fdfbe4","#f0c6c3","#df91a3","#d46780","#d1eeea"))
ggsave("Italian_areas_of_studies.png")

ggplot(data = way,
       aes(x = Where.in.Italy.are.you.studying.did.you.study., 
           fill = Where.are.you.from.)) + 
  geom_bar(position = "fill") +
  labs(x = "Italian areas",
       title  = "Italian areas of studies by origin and off-site student status",
       fill = "Continent or Italian areas of origin") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  scale_fill_manual(values=c("#798234","#a3ad62","#d0d3a2","#fdfbe4","#f0c6c3","#df91a3","#d46780","#d1eeea")) +
  facet_grid(.~Are.you.an.off.site.student.)
ggsave("Italian_areas_of_studies_by_origin.png")


ggplot(data = way,
       aes(x = Where.in.Italy.are.you.studying.did.you.study., 
           fill = Is.bioinformatics.the.main.focus.of.your.studies.)) + 
  geom_bar(position = "dodge") +
  labs(x = "Italian areas",
       title  = "Italian areas of studies by bioinformatics as main topic",
       fill = "Bioinformatics as main topic") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
   scale_fill_manual(values=c("#66C5CC","#DCB0F2"))
ggsave("Italian_areas_of_studies_by_bioinfo.png")

```

# What is your current position?

```{r,hide = TRUE}
ggplot(data = way,
       aes(x = What.is.your.current.position., 
           fill = What.is.your.current.position.)) + 
  geom_bar(aes(y = (..count..)/sum(..count..)),show.legend=FALSE) + 
  scale_y_continuous(labels=scales::percent) +
  labs(x = "Current position", 
       y = "Relative frequency", 
       title  = "What is your current position ?") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 50, hjust=1)) +
  scale_fill_manual(values=c("#245668","#04817E","#39AB7E","#8BD16D","#EDEF5D"))
ggsave("Current_position.png")


ggplot(data = way,
       aes(x = How.old.are.you., 
           fill = What.is.your.current.position.)) + 
  geom_bar(position = "dodge") +
  labs(x = "Age group",
       title  = "Current position by age group",
       fill = "Current position") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  scale_fill_manual(values=c("#245668","#04817E","#39AB7E","#8BD16D","#EDEF5D"))
ggsave("Current_position_by_age.png")
```

# What do non bioinfo participants study ?
```{r}
library(plyr)
library(ggplot2)
library(webr)
library(dplyr)
library(data.table)
way_study <- way %>% select("Is.bioinformatics.the.main.focus.of.your.studies.",
"What.is.your.current.degree.area..Please.select.the.closest.answer.that.applies..")

setDT(way_study)[, Freq := .N, by = .(Is.bioinformatics.the.main.focus.of.your.studies.,What.is.your.current.degree.area..Please.select.the.closest.answer.that.applies..)]
DF <- way_study %>% unique()

colnames(DF)[1] <- "Bioinformatics"
DF$What.is.your.current.degree.area..Please.select.the.closest.answer.that.applies.. <- revalue(DF$What.is.your.current.degree.area..Please.select.the.closest.answer.that.applies.., c(
"Life Sciences (Biology/Biotechnology)"="Life Science",
"Mathematics/Statistics"="Maths/Stats",
"Geo Sciences (Geology/Earth Sciences/Geophysics)"="Geo Sciences",
"Pharmacy/Pharmacology"="Pharma"))

#Pie Donut Chart:
PieDonut(DF, aes(Bioinformatics, What.is.your.current.degree.area..Please.select.the.closest.answer.that.applies.., count=Freq),showRatioThreshold = 0,donutLabelSize = 3,r1=0.7,r2=1,labelposition=1)
```

# Plots for poster

## Current position by bioinfo/not bionfo (donut chart)
```{r,hide = TRUE}
# load library
library(ggplot2)
library(ggpubr)
 
# Create dataset with category and count:
## split in bioinfo/no bioinfo :
way_bioinfo <- way %>% filter(Is.bioinformatics.the.main.focus.of.your.studies. == "Yes") %>% select(What.is.your.current.position.)
way_nobioinfo <- way  %>% filter(Is.bioinformatics.the.main.focus.of.your.studies. == "No") %>% select(What.is.your.current.position.)


# Function for donut_chart: 
# Compute percentages
donut_chart_fc <- function(data,title_name) {
    colnames(data)[1] <- "category"
data$category <- revalue(data$category, c(
"Bachelor’s student"="Bachelor student",
"Employed in other academic positions (research assistant, facility technician, …)"="Employed in other academic positions",
"Employed outside academia (company, start-up, spin-off, …)"="Employed outside academia",
"PhD student (either in academia or industry)"="PhD student"))
data <- data %>%
  group_by(category) %>%
  mutate(count = n()) %>% unique()

data$fraction = data$count / sum(data$count)

# Compute the cumulative percentages (top of each rectangle)
data$ymax = cumsum(data$fraction)

# Compute the bottom of each rectangle
data$ymin = c(0, head(data$ymax, n=-1))
 
# Compute label position
data$labelPosition <- (data$ymax + data$ymin) / 2

# Compute a good label
data$label <- paste0(round(data$fraction*100,digits = 2),"%")

# Make the plot
donut_chart_plot <- ggplot(data, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=category)) +
  geom_rect() +
  geom_label( x=3.57, aes(y=labelPosition, label=label), size=5) +
  scale_fill_brewer(palette=4) +
  coord_polar(theta="y") +
  xlim(c(2, 4)) +
  theme_void() +
  guides(fill = guide_legend(
    title = "Legend",
    override.aes = aes(label = ""))) +
  theme(legend.position = "left",legend.title = element_text(size = 12),plot.title = element_text(size = 20)) +
  ggtitle(title_name)
  
return(donut_chart_plot)
}


way_bioinfo_donut <- donut_chart_fc(way_bioinfo,"Bioinfo")
#need to delete legend in no bioinfo:
way_nobioinfo_donut <- donut_chart_fc(way_nobioinfo,"No Bioinfo")

#combine the two plots:
donut_combined <- ggarrange(
way_bioinfo_donut, way_nobioinfo_donut,
          ncol = 2, nrow = 1, common.legend = TRUE, legend="bottom")

ggsave("Current_position_bybioinfo.png", donut_combined, width = 16, height = 8)
```

## Add pie chart in the middle
```{r,hide = TRUE}
#Load ggplot2
library(ggplot2)
library(dplyr)
library(gridExtra)
library(grid)
library("ggpubr")

# Create Data
way_bioinfo_status <- way %>% select(Is.bioinformatics.the.main.focus.of.your.studies.)

data <- way_bioinfo_status
# Compute the position of labels
pie_chart_fc <- function(data){
  colnames(data)[1] <- "group"
  data$group <- revalue(data$group, c("Yes"="B","No"="!B"))
  
  data$group <- as.factor(data$group)
  data <- data %>%
  group_by(group) %>%
  mutate(value = n()) %>% unique()
  
  
  
  data <- data %>% 
  arrange(desc(group)) %>%
  mutate(prop = value / sum(data$value) *100) %>%
  mutate(ypos = cumsum(prop)- 0.5*prop )
  
  
  data$label <- paste0(data$group, "\n",round(data$prop,digits=0),"%")

# Basic piechart
piechart_plot <- ggplot(data, aes(x="", y=prop, fill=group)) +
  geom_col(color = "white") +
  geom_text(aes(label = label),
            position = position_stack(vjust = 0.5),
            size=7) +
  coord_polar(theta = "y") +
  scale_fill_brewer(palette=3) +
  theme_void() 
  # + theme(legend.position="none")


 return(piechart_plot)
}

pie_chart_bioinfo <- pie_chart_fc(way_bioinfo_status)

##remove title:
way_bioinfo_donut <- way_bioinfo_donut + labs(title=NULL)
way_nobioinfo_donut <- way_nobioinfo_donut + labs(title=NULL)


combined_donut_piechart <- ggarrange(way_nobioinfo_donut, pie_chart_bioinfo, way_bioinfo_donut,
          ncol = 3, nrow = 1, common.legend = TRUE, legend="bottom", labels = NULL, widths = c(1.5, 0.7, 1.5))
          

ggsave("poster_figs/current_position_bybionfo_piechart.png", combined_donut_piechart, width = 20, height = 8)
```

## Where do bioinfo and non-bioinfo study combined with off-site student (no, yes same area of origin, yes different area of origin)
```{r,hide = TRUE}
#create column: off_site_origin with levels: 'No','Yes same origin region', 'Yes different origin region'
#need to have same levels for the two columns I am comparing, so add levels to 'Where.in.Italy.are.you.studying.did.you.study'
way$Where.in.Italy.are.you.studying.did.you.study. <- factor(way$Where.in.Italy.are.you.studying.did.you.study., levels=levels(way$Where.are.you.from.))

way <- way %>% mutate(off_site_origin = ifelse(way$Are.you.an.off.site.student. == "No", "No", ifelse(way$Where.are.you.from. == way$Where.in.Italy.are.you.studying.did.you.study., "Yes same origin region", "Yes different origin region")))

way$off_site_origin <- as.factor(way$off_site_origin)

Italian_areas_study_by_offsiteorigin_bioinfostatus <- ggplot(data = way,
       aes(x = Where.in.Italy.are.you.studying.did.you.study., 
           fill = Where.in.Italy.are.you.studying.did.you.study.)) + 
  geom_bar(aes(y = (..count..)/sum(..count..)),show.legend=FALSE) + 
  scale_y_continuous(labels=scales::percent) +
  labs(x = "Italian areas of study",
       y = "Relative frequency",
       title  = "Italian areas of studies by bioinformatics/STEM and off-site status") +
  theme_classic() + 
  facet_grid(Is.bioinformatics.the.main.focus.of.your.studies. ~ off_site_origin) +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  scale_fill_manual(values=c("#245668","#04817E","#39AB7E","#8BD16D","#EDEF5D"))

ggsave("poster_figs/Italian_areas_study_by_offsiteorigin_bioinfostatus.png",Italian_areas_study_by_offsiteorigin_bioinfostatus, width = 16, height = 8)  
```
